\textsf{Are large language models superhuman chemists?}

### 2024-10-25 src/tex/appendix.tex
Update src/tex/appendix.tex
chore: update the format of the examples table (#74)
Co-authored-by: sreekanth221998 <sreekanthkuncha@gmail.com>
Co-authored-by: Kevin M Jablonka <32935233+kjappelbaum@users.noreply.github.com>

        Question Type & \% & Example Questions  \\

@@

\begin{xltabular}{\textwidth}{p{2.7cm}p{4em}X}
        \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question.} \\

        \toprule
        Required Skills & Number of Questions & Example Question \\
        \midrule
        Knowledge & 919 & Which of the following salts is an acidic salt? \\
        & & A. \ce{NH4Cl} \\
        & & B. \ce{Na2CO3} \\
        & & C. \ce{NaH2PO4} \\
        & & D. \ce{Zn(OH)Cl} \\
        \midrule
        Reasoning & 987 & How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
        & & A. By atomic force microscopy \\
        & & B. By transform infrared spectroscopy \\
        & & C. By measuring a photoluminescence spectrum \\
        & & D. By absorption spectroscopy \\
        \midrule
        Intuition & 1001 & Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
        & & A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
        & & B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
        \midrule
        Calculation & 128 & Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
        \bottomrule

    \label{tab:chembench_corpus_cognitive}
\end{xltabular}

>>>>

            \multicolumn{1}{c}{\textbf{Analytical Chemistry} 166 Questions (5.8\%)} \\

@@

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples for each of the required skills considered in the \chembench corpus.} The table shows the number of questions for each skill and an example question. Note that the total count in this table is bigger than the \chembench corpus. This is because the same question can be annotated with two different \enquote{requires} skills, i.e. Reasoning and Calculation.} \\
    \label{tab:chembench_corpus_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
            \midrule
            Which of the following salts is an acidic salt? \\
            A. \ce{NH4Cl} \\
            B. \ce{Na2CO3} \\
            C. \ce{NaH2PO4} \\
            D. \ce{Zn(OH)Cl} \\
            \midrule
            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            \midrule
            How can the optical impact of quantum confinement on the electronic structure of a quantum dot be observed experimentally? \\
            A. By atomic force microscopy \\
            B. By transform infrared spectroscopy \\
            C. By measuring a photoluminescence spectrum \\
            D. By absorption spectroscopy \\
            \midrule
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
            \midrule
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            A. [START\_SMILES]CC1(C)Oc2ccc([N+](=O)[O-])cc2[C@@H](N2CCOCC2)[C\@\@H]1O[END\_SMILES] \\
            B. [START\_SMILES]Cc1ccccc1N=C(S)N1CCC(NC\-(=O)c2ccco2)CC1[END\_SMILES] \\
            \midrule
            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
            \midrule
            Given that the average molar mass of the polymer chains in this sample of poly(lactic acid) (PLA) is \SI{595}{g mol^{-1}} using end-group analysis, where \SI{0.1619}{g} of PLA was dissolved in \SI{25}{cm^3} of benzyl alcohol and titrated with \SI{0.0400}{mol dm^{-3}} \ce{KOH} solution. The volume of \ce{KOH} solution required to reach the endpoint was \SI{6.81}{cm^3}. What is the average number of monomer units in each polymer chain of this sample? \\
            \bottomrule
        \end{tabularx}
    }
\end{table}

In analyzing the performance of state-of-the-art models on the ChemBench corpus, we identified several chemical topics where the models consistently struggled to provide correct answers. The following table \cref{tab:chembench_corpus_models_failed_topic} presents examples of challenging questions across different chemical domains, highlighting specific areas where current models need improvement

\begin{table}
    \centering
    \small
    \caption{\textbf{Examples of topics evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly.} The table displays the percentage of questions in the corpus for each topic, along with example questions and the percentage of questions that were not answered correctly by the models.} \\
    \label{tab:chembench_corpus_models_failed_topic}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \multicolumn{1}{c}{\textbf{Analytical Questions} 166 Questions (5.8\%) (26.5\%)} \\
            Which of the following methods can be used to support the structure elucidation of planar molecules by examining the topography of individual molecules? \\
            A. DESI-mass spectrometry \\
            B. XPS \\
            C. atomic force microscopy \\
            D. scanning tunneling microscopy \\
            E. secondary ion mass spectrometry \\
            \multicolumn{1}{c}{\textbf{Chemical Preference} 1001 Questions (35\%) (8.5\%)} \\
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            [START\_SMILES]N\#Cc1ccc(OCCCN2CC3CN\-(CCNS(=O)(=O)c4ccc(F)cc4)CC(C2)O3)cc1\-[END\-\_SMILES] \\
            [START\_SMILES]O=C1CC(c2ccc(CC(NS(=O)\-(=O)\-c3cc(Cl)cc(Cl)c3)c3nc4ccccc4[nH]3)cc2)S\-(=O)\-(=O)N1[END\_SMILES] \\
            \multicolumn{1}{c}{\textbf{General Chemistry} 152 Questions (5.3\%) (3.2\%)} \\
            Why \ce{SCl6} does not exist if other similar molecules such as \ce{SF6} exist? \\
            A. \ce{SCl6} does not exist because of steric effects \\
            B. \ce{SCl6} does not exist because of the high electronegativity of chlorine \\
            C. \ce{SCl6} does not exist because of the high ionization potential of the chlorine \\
            D. \ce{SF6} exist because of the high electronegativity of fluorine \\
            \multicolumn{1}{c}{\textbf{Inorganic Chemistry} 94 Questions (3.3\%) (4.2\%)} \\
            What is the spin-only magnetic moment (in units of Bohr-magnetons) for the low-spin case of the metal complex \ce{RhCl(CO)(PPh3)2}?\\
            \multicolumn{1}{c}{\textbf{Materials Science} 89 Questions (3.1\%) (0.9\%)} \\
            What are the conditions for the observation of effects of quantum confinement in quantum dots? \\
            A. All dimension of the material must be smaller than exciton Bohr radius \\
            B. One dimension of the quantum dot must be smaller than exciton Bohr radius \\
            C. The movement od charge carriers is limited in all spatial dimensions \\
            \multicolumn{1}{c}{\textbf{Organic Chemistry} 441 Questions (15.5\%) (0.5\%)} \\
            What makes acrylates more reactive than typical hydrocarbons? \\
            A. The carbonyl group \\
            B. The double bond \\
            C. The hydroxyl group \\
            D. The vinyl group \\
            \multicolumn{1}{c}{\textbf{Physical Chemistry} 179 Questions (6.3\%) (13.4\%)} \\
            Why are polar molecules more soluble in water than non-polar molecules? \\
            A. Non-polar molecules should break hydrogen bonds in water to dissolve \\
            B. The entropy of the system increases when non-polar molecules dissolve in water \\
            C. Water has a dipole and can easier isolate polar molecules from each other \\
            D. Water makes stable dipole-dipole interactions with polar molecules \\
            \multicolumn{1}{c}{\textbf{Technical Chemistry} 42 Questions (1.5\%) (4.7\%)} \\
            Which of the following statements is/are true about granulation shaping of solid catalysts? \\
            A. Broader product diameter distribution than pellets or extrudates \\
            B. Facilitates reactor upscaling \\
            C. It is also called snow-ball \\
            D. Not suitable to prepare egg-shell catalysts \\
            \multicolumn{1}{c}{\textbf{Toxicity/Safety} 690 Questions (24.2\%) (40.4\%)} \\
            What determines the flammability of polyols? \\
            A. Their chemical structure \\
            B. Their low volatility \\
            C. Their odor threshold \\
            D. Their reaction with acids \\
        \end{tabularx}
    }
\end{table}

To better understand the cognitive demands of chemical assessment, we categorized the questions in the ChemBench corpus according to the primary skills required to solve them. The table \Cref{tab:chembench_corpus_models_failed_cognitive} below shows representative examples of questions that state-of-the-art models failed to answer correctly across different cognitive skill categories, including knowledge, reasoning and intuition.
\begin{table}
    \centering
    \small
    \caption{\textbf{Examples of required skills evaluated in the \chembench corpus that the state-of-the-art model failed to answer correctly} The table displays the number of questions for each skill, along with an example question that the model failed to answer.} \\
    \label{tab:chembench_corpus_models_failed_cognitive}
    {\fontsize{8}{9}\selectfont
        \begin{tabularx}{\textwidth}{X}
            \toprule
            \multicolumn{1}{c}{\textbf{Knowledge} 919 Questions} \\
            \midrule
            Which of the following are true for phosphoric acid? \\
            A. Is a weak acid, so it is not toxic \\
            B. Should never be tasted \\
            C. Though a component of soft drinks, reacts vigorously with water when in concentrated form \\
            \midrule
            \multicolumn{1}{c}{\textbf{Reasoning} 987 Questions} \\
            \midrule
            Which of the following elements has the largest atomic radius? \\
            A. \ce{Fr} \\
            B. \ce{He} \\
            C. \ce{Li} \\
            D. \ce{Rb} \\
            \midrule
            \multicolumn{1}{c}{\textbf{Intuition} 1001 Questions} \\
            \midrule
            Imagine an early virtual screening campaign setting (accounting for simple aspects such as oral availability and small molecular profile, but no other modalities such as covalency or bifunctionality). Which of the following two candidates would you prefer for further development? \\
            A. [START_SMILES]CCc1c(CCOC)nn(-c2ccccc2)c1-c1ccccc1[END_SMILES] \\
            B. [START_SMILES]Cc1cc(/C=C2\\C(=N)N3N=C(S(C)(=O)=O)SC3=NC2=O)c(C)n1Cc1ccccc1[END_SMILES] \\
            \midrule
            \multicolumn{1}{c}{\textbf{Calculation} 128 Questions} \\
            \midrule
            The cell potential for the overall cell reaction during electrolysis of water, \ce{2H2O(l) -> 2H2(g) + O2(g)} is \SI{-1.23 V}. What is the standard electrode potential for the half reaction, \ce{2H2O(l) -> O2(g) + 4H+(aq) + 4e-} in V? \\
            \bottomrule
        \end{tabularx}
    }
\end{table}
